A Large-scale Study of Polymorphism in R

### 2019-03-27 paper/main.tex
minor
typos, behaviour->behavior, updated corpus counts

\newcommand{\PACKAGES}{7,763\xspace}
\newcommand{\PROGRAMMERS}{10,000\xspace}
\newcommand{\PERCENTCRAN}{56\%\xspace}
\newcommand{\CRANTOTAL}{13,841\xspace}
\newcommand{\RLOC}{8,996,572\xspace}
\newcommand{\CLOC}{5,100,719\xspace}
\newcommand{\YEARS}{20\xspace}

@@

R and it's ancestor S are languages designed, implemented and maintained by
statisticians. Originally they were designed as glue languages, languages
that would allow to read data into vectors and call statistical routines
written in Fortran. Over three decades, the languages became widely used
across many fields of science and in industry for data analysis and data
visualization; with time additional features were added.  Modern R, as a
linguistic object of study, is fascinating. It is a vectorized, dynamically
typed, lazy functional language with limited side-effects, extensive
reflective facilities and retrofitted object-oriented programming support.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{linesofrandccode FILE MISSING IN GIT}
\caption{Lines of code in the analyzed corpus. \# of lines of R code are above 0, \# of lines of C code are below 0.}
\label{allcloc}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.1]{recordsbypkg}
\caption{Numbers of Recorded Function Invocations in the Analyzed Corpus}
\label{recorded}
\end{center}
\end{figure}

@@

These naturally-occurring attribute patterns and classes can be separately
accounted for to highlight user-defined patterns and behaviours.  This will
be made clear in Section~\ref{sec:results}.

@@

In Figure~\ref{fig:matvec}, we see that the function can take in a vector, but immediately transforms it (transpose, with {\tt t}) into a matrix.

@@

In terms of programmer behaviour, we observe a pattern of optional arguments throughout several R packages, wherein programmers specify a default argument value which is sometimes NULL.
Consider the snippet in Figure~\ref{fig:optnull} for an example.
\AT{TODO put that figure here}
The distinction between NULL and other types is not particularly useful to the programmer, so we propose to also collapse this distinction.

>>>>

\newcommand{\PACKAGES}{11,463\xspace}
\newcommand{\PROGRAMMERS}{?\xspace}
\newcommand{\PERCENTCRAN}{83\%\xspace}
\newcommand{\CRANTOTAL}{13,841\xspace}
\newcommand{\RLOC}{15,050,267\xspace}
\newcommand{\CLOC}{9,373,542\xspace}
\newcommand{\YEARS}{?\xspace}

@@

R and its ancestor S are languages designed, implemented and maintained by
statisticians. Originally they were designed as glue languages, languages
that would allow to read data into vectors and call statistical routines
written in Fortran. Over three decades, the languages became widely used
across many fields of science and in industry for data analysis and data
visualization; with time additional features were added.  Modern R, as a
linguistic object of study, is fascinating. It is a vectorized, dynamically
typed, lazy functional language with limited side-effects, extensive
reflective facilities and retrofitted object-oriented programming support.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{linesofrandccode}
\caption{Lines of code in the analyzed corpus. \# of lines of R code are above 0, \# of lines of C code are below 0.}
\label{allcloc}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.6]{recordsbypkg}
\caption{Numbers of Recorded Function Invocations in the Analyzed Corpus}
\label{recorded}
\end{center}
\end{figure}

@@

These naturally-occurring attribute patterns and classes can be separately
accounted for to highlight user-defined patterns and behaviors.  This will
be made clear in Section~\ref{sec:results}.

@@

In Figure~\ref{fig:matvec}, we see that the function can take in a vector, but immediately transforms it (transpose, with {\tt t}) into a matrix.
As an idea, matrix/vector polymorphism seems sensible, as mathematically vectors are matrices.
R echoes this by making conversion between the two ``types'' easy:
\code{as.vector(m)} flattens a matrix \code{m} into a vector (e.g., \code{as.vector(matrix(2, 2, 2)) == c(2, 2, 2, 2)}), and \code{as.matrix(v)} builds a {\tt length(v) x 1} matrix (e.g., \code{as.matrix(c(1, 2)) == matrix(data=c(1, 2))}).
\AT{TODO fix line break in code at end of last sentence}

@@

In terms of programmer behavior, we observe a pattern of optional arguments throughout several R packages, wherein programmers specify a default argument value which is sometimes NULL.
Consider the snippet in Figure~\ref{fig:optnull} for an example.
\AT{TODO put that figure here}
The distinction between NULL and other types is not particularly useful to the programmer, so we propose to also collapse this distinction.
